---
title: "Case Study"
author: "Bhavikkumar Mistry"
date: "2024-06-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*About a company*
Bellabeat, a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company.

*Ask*
**Questions for Analysis**
-What are some trends in smart device usage?
-How could these trends apply to Bellabeat customers?
-How could these trends help influence Bellabeat marketing strategy



*Prepare*
**Dataset**

The dataset used in this analysis was obtained from Fitbit Fitness Trackers and accessed through Kaggle public domain FitBit Dataset (CC0). The data was collected from thirty-three eligible Fitbit users who consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep through Amazon Mechanical Turk between April 12, 2016 and May 12, 2016. The dataset is updated annually with the most recent being on March 2nd, 2024. The data consists of 18 CSV files organized in long format. Each document represents different quantitative data tracked by the user's Fitbit. Every user has a unique ID and the data was tracked by either by day, hour or minute. Given the small sample size, the data was initially analyzed using Microsoft Excel. The relations between the tables were validated and the attributes and length of the survey were summarized using pivot tables. The Id field was used as the key field to merge the data sets.

**Data Privacy**

This Kaggle data set contains personal fitness tracker data from thirty-three Fitbit users. The users consented to the submission of personal tracker data and this is made available to the public under a CC0 license.

**Data Bias and Limitations**

The sample size of thirty-three individuals is very small and therefore not representative of the overall population. Also, there is no demographic information associated with the data. Given that Bellabeat’s main users are women, knowing the gender of the sample would be advantageous for the analysis and subsequent recommendations. Some individual logs, such as the weight log, are particularly under sampled and only contain data from eight users; two-thirds of the entries were manually entered. Additionally, many logs contain null values. The data is eight years old and outdated. It is, however, cited many times on Kaggle and updated frequently.

**Additional Data Sets**

Given the small sample size of the data set an additional data set could be used to increase the confidence level, leading to more robust conclusions. Another approach would be to use a data set that considers individuals over an extended period of time. One such data set comes from the Mi Band fitness tracker data. It contains data relating to steps and sleep for one individual from April 2016 to July 2022 collected from the Mi Band.

*Process*

I chose RStudio to clean and prepare my data as it can handle vast amount of dataset and is versitile for Data sharing in form of Markdowns which will come handy in Share phase.
Also It will help me document each step of the process when I clean my data.

```{r}
#First import some packages

library(tidyverse) 
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)

```

Importing Datasets
```{r}
activity <- read.csv("../bellaboom/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
calories <- read.csv("../bellaboom/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/hourlyCalories_merged.csv")
intensities <- read.csv("../bellaboom/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/hourlyIntensities_merged.csv")
sleep <- read.csv("../bellaboom/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
weight <- read.csv("../bellaboom/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/weightLogInfo_merged.csv")
steps <- read_csv("../bellaboom/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailySteps_merged.csv")
```
```{r, echo=FALSE}
head(steps
     , 10)
```
Fixing timestamps 
```{r}
#Intensity
intensities$ActivityHour = as.POSIXct(intensities$ActivityHour, format ="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
intensities$time <- format(intensities$ActivityHour, format = "%H:%M:%S")
intensities$date <- format(intensities$ActivityHour, format = "%m/%d/%y")
# calories
calories$ActivityHour=as.POSIXct(calories$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
calories$time <- format(calories$ActivityHour, format = "%H:%M:%S")
calories$date <- format(calories$ActivityHour, format = "%m/%d/%y")
# activity
activity$ActivityDate=as.POSIXct(activity$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
activity$date <- format(activity$ActivityDate, format = "%m/%d/%y")
# sleep
sleep$SleepDay=as.POSIXct(sleep$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
sleep$date <- format(sleep$SleepDay, format = "%m/%d/%y")

```
**Data Exploration**
```{r}
n_distinct(activity$Id)
n_distinct(calories$Id)
n_distinct(intensities$Id)
n_distinct(sleep$Id)
n_distinct(weight$Id)
n_distinct(steps$Id)
```
There are very 33 different individual's data is present. The sleep data only has 24 individuals. However Weight has only 8 individuals which is not enough data to make any solid conclusion.
Lets Explore more each dataset
```{r}
# Data cleaning - checking for duplicate entries

sum(duplicated(activity))
sum(duplicated(steps))
sum(duplicated(sleep))
sum(duplicated(intensities))
sum(duplicated(steps))
```
```{r}
# Data cleaning - remove duplicates and drop rows with missing values in any columns



sleep <- sleep %>%
  distinct() %>%
  drop_na()

#verify missing data
sum(duplicated(sleep))

```
*Analyse*
```{r}
#Activity summary of sedentary time
activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes, Calories) %>%
  summary()

#Acrivity summary of active time
activity %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes) %>%
  summary()

#Calories
summary(calories) 

# sleep
sleep %>%
  select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>%
  summary()

# weight
weight %>%
  select(WeightKg, BMI) %>%
  summary()

#steps
  summary(steps)
```

**Few intresting discoveries**
-Average Step count is 7638 per day. The Centers for Disease Control and Prevention (CDC) recommends walking at least 10,000 steps per day.
-Average Sedentary time is 991 minutes which is 16 hours/day. Which is not ideal, users must be made aware of excercising.
-There are average of 1 sleeping session with 6.8 hours of asleep time.However, avg time in bed is 7.6 hours, some people might have difficulties fall asleep or due to some other reasons.We can send notifications to users to sleep in that avg 0.8 hours (58 minutes of unsleepyness)
-People in majority are lightly active.

**Merging Data**
```{r}

merged_data <- merge(sleep, activity, by=c('Id', 'date'))
head(merged_data)
```


*Share & Act*
Lets Plot some Relationships
```{r, warning=FALSE}
# Analysis - relationship between steps taken in a day and sedentary minutes
ggplot(data=merged_data, aes(x=TotalSteps, y=SedentaryMinutes))+ geom_point() + 
          labs(title = "Daily Steps vs Sedentary Time")


#Relation between TimeAsleep vs TimeinBed
ggplot(data=merged_data, aes(x=TotalMinutesAsleep, y=SedentaryMinutes))+geom_point() +
          geom_smooth() + labs(title = "Time Asleep vs Time in Bed")

#Realtion in steps and claories burned
ggplot(data = merged_data, aes(x=Calories, y=TotalSteps, group=date))+geom_point() + labs(title= "Steps vs Calories burned")

#Intentensity breakdown by time
int_new <- intensities %>%
  group_by(time) %>%
  drop_na() %>%
  summarise(mean_total_int = mean(TotalIntensity))

ggplot(data=int_new, aes(x=time, y=mean_total_int)) + geom_histogram(stat = "identity", fill='green') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Average Total Intensity vs. Time")

```

Observations:
-There is no relation between Daily steps to sedentary time.
-however, Time asleep decreases with More sedentary minutes. We can alert users to be more active and reduce their sedentary minutes in order to improve their sleep.
-There is relation between Total steps and Calories burned but not strong. There might be other activities that might be accountable for calories burned.
-The Intensity between 5 pm to 7 pm are highest.This also some of the busiest times people hit the Gym. We can remind bellabeat app users to workout.
***Keep in mind that we need to support this insights with more data, because correlation between some data doesn’t mean causation.***

